From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-21 dictionary.tex
More remarks and examples.

The main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on reductive groups
over algebraically closed fields with finite cyclic component groups in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, such groups are extensions of finite cyclic groups by algebraic tori,
in which case it is not difficult to relate Frobenius-stable character sheaves to our character sheaves (\S \ref{ssec:alg_groups}). The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ do not arise in that context.

@@

\section{Recollections for connected algebraic groups}\label{ssec:connected}

@@

\begin{proposition} \label{prop:etale-iso}
 If $G$ is \'etale then the map $S_G : \CSiso{G}\to \Hh^2(E^\bullet_G)$ induces an isomorphism of split short exact sequences
\[
\begin{tikzcd}
 0 \arrow{r} & \ker \TrFrob{G} \arrow{d} \arrow{r} & \CSiso{G}\arrow{d}{S_G} \arrow{r}{\TrFrob{G}} \arrow{r} & G(\Fq)^* \arrow{d} \arrow{r} & 0\\
  0 \arrow{r} & \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \arrow{r} & \Hh^2(E^\bullet_G) \arrow{r} & \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \arrow{r} & 0.
 \end{tikzcd}
 \]
\end{proposition}
\begin{proof}
This result follows from Propositions~\ref{prop:ses_desc}, \ref{prop:SGiso} and \ref{prop:sur_etale}.
\end{proof}

@@

\begin{proposition}\label{prop:autornaught_etale}
Let $G$ be an \'etale commutative group scheme over $\Fq$.
If $\cs{L}$ and $\cs{L}'$ are character sheaves on $G$ then
every $\rho\in \Hom(\cs{L},\cs{L}')$ is either trivial or an isomorphism. Moreover, the trace map induces an isomorphism of groups
\[
\Aut_\CS{}(\cs{L}) \to \Hom(G(\bFq)_{\Weil{}}, \EEx).
\]
\end{proposition}

@@

\begin{example}
Consider the case when $G$ is the discrete \'etale group scheme $\ZZ$.
If $\chi : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi(1) = \ell$
and if $\cs{L}$ is a character sheaf on $G$ in the isomorphism class
corresponding to $\chi$ under Theorem~\ref{thm:snake},
then $\cs{L}$ does not descend to $G$ since the image of $\chi$ is not
bounded.
On the other hand, if  the character $\chi : \ZZ \to \EEx$ of $G(\Fq)$ determined by $\chi(1) = 1+\ell$.
The stalk of $\cs{L}$ at the component $1$ is defined
by a pro-\'etale cover  . . . 
\end{example}

@@

\begin{corollary}\label{cor:alg_groups}
Let $G$ be a commutative algebraic group over $\Fq$.
All character sheaves on $G$ descend to $G$: $\CS(G)$ is equivalent to $\bCS(G)$ (\ref{ssec:bS}).
The trace of Frobenius determines a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \bCSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
The group $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ need not be trivial.
\end{corollary}

@@

Add comment relating to Lusztig . . . .

>>>>

The main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on certain reductive groups over algebraically closed fields in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, it is not difficult to relate Frobenius-stable character sheaves to our character sheaves (Remark~\ref{rem:Lusztig}). The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ do not arise in that context.

@@

\section{Recollections on character sheaves for connected algebraic groups}\label{ssec:connected}

@@

\begin{proposition} \label{prop:etale-iso}
 If $G$ is \'etale then the map $S_G : \CSiso{G}\to \Hh^2(E^\bullet_G)$ induces an isomorphism of short exact sequences
\[
\begin{tikzcd}
 0 \arrow{r} & \ker \TrFrob{G} \arrow{d} \arrow{r} & \CSiso{G}\arrow{d}{S_G} \arrow{r}{\TrFrob{G}} \arrow{r} & G(\Fq)^* \arrow{d} \arrow{r} & 0\\
  0 \arrow{r} & \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \arrow{r} & \Hh^2(E^\bullet_G) \arrow{r} & \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \arrow{r} & 0.
 \end{tikzcd}
 \]
\end{proposition}
\begin{proof}
This result follows from Propositions~\ref{prop:ses_desc}, \ref{prop:SGiso} and \ref{prop:sur_etale}.
\end{proof}

@@

\begin{proposition}\label{prop:autornaught_etale}
Let $G$ be an \'etale commutative group scheme over $\Fq$.
If $\cs{L}$ and $\cs{L}'$ are character sheaves on $G$ then
every $\rho\in \Hom(\cs{L},\cs{L}')$ is either trivial or an isomorphism. Moreover, the trace map induces an isomorphism of groups
\[
\Aut(\cs{L}) \to \Hom(G(\bFq)_{\Weil{}}, \EEx).
\]
\end{proposition}

@@

\begin{example}
Consider the case when $G$ is the discrete \'etale group scheme $\ZZ$.
If $\chi : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi(1) = \ell$
and if $\cs{L}$ is a character sheaf on $G$ in the isomorphism class
corresponding to $\chi$ under Theorem~\ref{thm:snake},
then $\cs{L}$ does not descend to $G$, since the image of $\chi$ is not
bounded.
If $\chi' : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi'(1) = 1+\ell$ and if $\cs{L}'$ corresponds to $\chi'$ under Theorem~\ref{thm:snake},
then $\cs{L}'$ does descend to $G$, since the image of $\chi'$ is
bounded.
However, $\cs{L}'$ is not defined by a discrete isogeny (\S \ref{ssec:discrete_isogenies}). 
If $\chi'' : \ZZ \to \EEx$ is the character of $G(\Fq)$ determined by $\chi''(1) = \zeta$, a root of unity in $\EEx$, and if $\cs{L}''$ corresponds to $\chi''$ under Theorem~\ref{thm:snake},
then $\cs{L}''$ is defined by a discrete isogeny. 
\end{example}

@@

\begin{corollary}\label{cor:alg_groups}
Let $G$ be a commutative algebraic group over $\Fq$.
All character sheaves on $G$ descend to $G$: $\CS(G)$ is equivalent to $\bCS(G)$ (\S \ref{ssec:bS}).
The trace of Frobenius determines a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \bCSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
The group $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ need not be trivial.
\end{corollary}

@@

\begin{remark}\label{rem:Lusztig}
Suppose $G$ is a commutative algebraic group over $\Fq$ appearing in the series of papers starting with \cite{lusztig:disconnected1}.
Then $G$ is a commutative reductive algebraic group over $\Fq$ with cyclic component group scheme; such groups are extensions of finite cyclic groups by algebraic tori.
Every Frobenius-stable character sheaf on $G$, in the sense of \cite{lusztig:disconnected1}, is a character sheaf on $G$, in our sense,
and vice versa.
Moreover, since $\Hh^2(\pi_0(\bG),\EEx)=0$ by Remark~\ref{eg:H2}, it follows that each Frobenius-stable character sheaf on $G$ as in \cite{lusztig:disconnected1}, is determined by its trace of Frobenius, up to isomorphism. 
\end{remark}
